---
title: "Multivariate Analisys Report"
subtitle: "(MECD, Minor-CD, & MMAC, 1st Semester, 2023/2024)"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2023-12-05"
header-includes:
  - \usepackage{titling}
  - \pretitle{\begin{center}
    \includegraphics[width=2in,height=2in]{IST_logo.png}\LARGE\\}
  - \posttitle{\end{center}}
---

**Professor: M. Rosário Oliveira**

**Group Members:**

| Antonio Villas Bôas | Darlan Nascimento | Emma Dennis-Knieriem | Leonardo Aikawa | Tunahan Güneş |
|:-:|:-:|:-:|:-:|:-:|
| ist1105429 | ist1108076 | ist1105545 | ist1105545 | ist1108108 |

# Fitness Club Dataset for ML Classification {style="align:center"}


\newpage

## Introduction

Multivariate analysis is a pivotal statistical technique in the realm of data science, particularly when dealing with datasets encompassing multiple interrelated variables. It offers a sophisticated approach to understanding the nuanced relationships and patterns that emerge from complex data. This method is crucial for extracting meaningful insights from datasets where multiple factors interact in intricate ways, as is often the case in real-world scenarios.

Multivariate analysis includes a variety of techniques, each suited to different types of data and analysis objectives. Key methods such as logistic regression, decision trees, and cluster analysis are particularly relevant in scenarios where predicting outcomes or classifying data points are the main goals. These techniques allow for a comprehensive analysis, considering how various factors collectively influence outcomes.

## Description of the problem under study

Our dataset was select from [Kaggle](https://www.kaggle.com/datasets/ddosad/datacamps-data-science-associate-certification/data). The dataset provided for this analysis originates from GoalZone, a well-known fitness club chain in Canada. GoalZone operates several fitness classes, with capacities of either 25 or 15 participants. A significant challenge they face is the discrepancy between class bookings and actual attendance. While some classes are consistently fully booked, the actual attendance rate often falls short of expectations. This scenario presents both a logistical challenge and an opportunity for optimization.

The primary challenge is predicting whether a member who has booked a class will actually attend. Accurately forecasting attendance allows GoalZone to optimize class capacities effectively, ensuring maximum utilization of resources and enhanced member satisfaction. This challenge involves analyzing various factors that might influence attendance, such as the time of the class, member demographics, previous attendance patterns, and class type.

## Objectives

The analysis of the GoalZone dataset aims to achieve the following objectives:

Predictive Analysis: Develop a predictive model that can accurately forecast whether a member who has booked a class will attend. This involves identifying key variables that influence attendance and employing appropriate multivariate techniques.
Optimization of Class Capacities: Use the insights gained from the predictive model to optimize class capacities, enabling GoalZone to allocate spaces more effectively and potentially increase the total number of spaces offered.
Enhancing Member Experience: By improving the accuracy of attendance predictions, GoalZone can reduce overbooking and under-utilization issues, leading to a better overall experience for members.
Data-Driven Decision Making: Provide GoalZone with data-driven strategies to manage class schedules and capacities, thereby improving operational efficiency.

You can also embed plots, for example:

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
# library(ggplot2)
```

```{r exploratory}
fitdata <- read_csv("fitness_class_2212.csv", show_col_types = FALSE)
str(fitdata)
summary(fitdata)
```

## Data Exploration and Cleaning

We noticed that some columns are not in the ideal format to work with our classification models. The column `days_before` should be a discrete (integer) number of days before the class the member registered, but in `fitdata` we got `class(fitdata$days_before)` as `r class(fitdata$days_before)`. Here are the steps we used to convert to integer:

```{r}
unique(fitdata$days_before)

fitdata$days_before <- fitdata$days_before %>%
     strsplit(split = " ") %>% # split some strings that had more than the number of days
     sapply(function(x) x[1]) %>% # gets the 1st element of the split strings
     as.integer() # converts to integer

class(fitdata$days_before)
```

For `day_of_week`, we needed to standardize to uniquely 7 values corresponding to the days of the week.
```{r}
unique(fitdata$day_of_week)

fitdata$day_of_week <- fitdata$day_of_week %>%
     tolower() %>% # converts to lower case
     substr(1, 3) %>% # gets only the 3 first elements of each string
     factor() # Converts to factor to be used as categorical variable

table(fitdata$day_of_week)
```





